---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data


```{r, echo = TRUE}

unzip("activity.zip")
activity <- read.csv("activity.csv")
```

## What is mean total number of steps taken per day?

```{r, echo = TRUE}
q1 <- aggregate(activity[,1],list(activity$date),sum,na.rm = TRUE)
hist(q1[,2], main = "Total number of step taken per day", xlab = "# of Steps", ylab = "# of Days", col = "grey")
```

The mean total number of steps taken per day is `r mean(q1[,2])`, the medium total number of steps taken per day is `r median(q1[,2])`.

## What is the average daily activity pattern?

```{r, echo = TRUE}
q2 <- aggregate(activity[,1],list(activity$interval), mean, na.rm = TRUE)
q2m <- max(q2[,2])
plot(x=q2[,1], y=q2[,2], type = "l", main = "Average Daily Activity Pattern", xlab = "Interval", ylab = "Average Steps")
```

The interval, on average across all the days in the dataset, contains the maximum number of steps is `r q2[q2$x == q2m,1]`.

## Imputing missing values
Total number of NAs is `sum(is.na(activity$steps))`.
 

```{r, echo = TRUE}
##My strategy of filling in the NAs is using the average steps taken for that specific interval.
##Create a new column in activity called nsteps that fills in all the NA with the average of that interval.
names(q2) <- c("interval","average")
nactivity <- activity
nactivity$nsteps <- ifelse(is.na(nactivity$steps), q2[q2$interval == nactivity$interval,2], nactivity$steps)
q3 <- aggregate(nactivity$nsteps,list(nactivity$date),sum)
hist(q3[,2], main = "Total number of step taken per day", xlab = "# of Steps", ylab = "# of Days", col = "dark grey")
```

The mean total number of steps taken per day is `r mean(q3[,2])`, the medium total number of steps taken per day is `r median(q3[,2])`. This mean and median is different from what we had in question one. Inputting missing data made the result more reliable. 

## Are there differences in activity patterns between weekdays and weekends?
```{r, echo = TRUE}
library(lattice)
## Create a column with the weekday info using "weekdays" function. Catogorize Sunday and Saturday as weekends and Monday through Friday as weekdays.
Sys.setlocale("LC_TIME", "English")
nactivity$temp <- weekdays(as.Date(nactivity$date), abbreviate = TRUE)
nactivity$weekdays <- ifelse(nactivity$temp %in% c("Sun","Sat"), "Weekend","Weekday")

##Calculate the average steps taken per interval.
q4 <- aggregate(nsteps~interval+weekdays, data = nactivity, mean)
xyplot(nsteps~interval | weekdays, data = q4 ,type = "l", layout = c(1,2))

```
